---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

 *1. Load the data (i.e. read.csv())*
```{r, echo=TRUE}
activity<-read.csv("activity.csv")
```

  *2. Process/transform the data (if necessary) into a format suitable for your analysis*
```{r, echo=TRUE}
complet_case<-complete.cases(activity)
comp_activity<-activity[complet_case==TRUE,]
```


## What is mean total number of steps taken per day?

  *1. Calculate the total number of steps taken per day*
```{r, echo=TRUE}
daily_steps<-split(comp_activity$steps,comp_activity$date)
daily_tot_steps<-sapply(daily_steps,sum)# getting the daily total number of steps
daily_tot_steps
```

  *2. Make a histogram of the total number of steps taken each day*
```{r, echo=TRUE}
hist(daily_tot_steps, main="Total Steps per Day", xlab="Steps") #Drawing the histogram and adding mean line and median line
abline(v=mean(daily_tot_steps), lty=2, col="red")               
abline(v=median(daily_tot_steps), lty=4, col="green")
```

  *3. Calculate and report the mean and median of the total number of steps taken per day*
```{r, echo=TRUE}
summary(daily_tot_steps) #gives the mean and the median
```


## What is the average daily activity pattern?

  *1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)*
```{r, echo=TRUE}
mean_steps_per_interval<-sapply(split(comp_activity$steps,comp_activity$interval),mean) #Getting the meean step per interval after ignoring nas
plot(mean_steps_per_interval,main="Mean Steps per Interval", xlab="Intervals")
abline(v=which.max(mean_steps_per_interval), lty=2, col="red")# Showing the maximum  
```

  *2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?*
```{r, echo=TRUE}

maximum<-which.max(mean_steps_per_interval)#retrieving its index
interval_max<-names(maximum)
interval_max #giving the interval with the max mean
mean_steps_per_interval[interval_max] #maximum value
```

## Imputing missing values

  *1.Calculate and report the total number of missing values in the dataset*
```{r, echo=TRUE}
nas<-which(is.na(activity$steps)==TRUE)#Getting indexes where value equal na
length(nas) #number of nas
```

  *2. Devise a strategy for filling in all of the missing values in the dataset*
```{r, echo=TRUE}
#Splitting the sreps per interval
list_steps_per_interval<-split(activity$steps,activity$interval)
#Replacing each na step value by the mean of it's belonging iterval
rep_nas_mean<- lapply(list_steps_per_interval, function(x) { 
  x[is.na(x)] <- mean(x, na.rm = TRUE)
  x
})
```

  *3. Create a new dataset that is equal to the original dataset but with the missing data filled in.*
```{r, echo=TRUE}
order_activity_per_interval<-activity[order(activity$interval,decreasing=FALSE),] # Ordering the initial dataset per interval
complete_dataset<-cbind("steps"=unlist(rep_nas_mean),order_activity_per_interval[,2:3]) # Appending the list of replaced steps_value to the two other columns of the initial dataset with no nas and ordered by interval
```

  *4. Make a histogram of the total number of steps taken each day*
```{r, echo=TRUE}
new_daily_tot_steps<-sapply(split(complete_dataset$steps,complete_dataset$date), sum)# getting the daily total number of steps with the new dataset
hist(new_daily_tot_steps, main="Total Steps per Day", xlab="Steps") #Drawing the histogram and adding mean line and median line
abline(v=mean(new_daily_tot_steps), lty=2, col="red")               
abline(v=median(new_daily_tot_steps), lty=4, col="green") 
```

  *Calculate and report the mean and median total number of steps taken per day*
```{r, echo=TRUE}
summary(new_daily_tot_steps)#Gives the mean and the median

```
Conclusion:   These values are a little bit different from the values we had in the first part of the assignement, because of the nas we had.

## Are there differences in activity patterns between weekdays and weekends?

  *1.  Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day*
  
```{r, echo=TRUE}
isWeekday<-lubridate::wday(complete_dataset$date) %in% c(1:5) #Telling whether a date is week day or not
complete_dataset<-cbind(complete_dataset,"levels"=factor(isWeekday,labels=c("weekend","weekday"))) #dataset with factor levels : weekday and weekend

```

  *2Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).*
```{r, echo=TRUE}
agg_complete_dataset<-aggregate(complete_dataset$steps, by=list(complete_dataset$interval, complete_dataset$levels), mean)
names(agg_complete_dataset)<-c("interval","levels","mean_steps") #adding header to the aggregate_variable
require(lattice)
xyplot(mean_steps ~ interval| levels, data = agg_complete_dataset, layout = c(1, 2),type='b') #Drawing the two Graphics
```


